Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2021-05-11 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

\begin{abstract} 
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition. Nevertheless, a general approach to envelope detection doesn't exist. Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation. To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning. The method draws inspiration from geometric concepts: the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope. Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave. The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible. We provide visualizations of the envelope extracted for various real-world signals with diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes. An optimized implementation is available via the Python Package Index (PyPI), while interactive visualizations and source code are available from the accompanying GitHub repository.
\end{abstract}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00AlphaShape.pdf}
  \caption{The alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that doesn't contain any point of the set, is also shown.}
  \label{fig:AlphaShape}
\end{figure}

@@

While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}, without intersecting with it.

@@

Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempirical}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.

@@

Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in representing those points on the Cartesian coordinate system, and for that we need a transformation that places those points on the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.

@@

In this way, we can consider both axes as having unit $ i $. Dividing them by $ i $, we obtain the effect of representing the points on the standard basis. Numerically, the relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points on the original coordinates later.

Considering $ P_{\mathbf{b}} $, a point of the set P represented on the original coordinate system and $ P_{\mathbf{s}} $ the same point represented on the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:

\begin{align}\label{eq:Basis}
\begin{split}
P & = (x, y) \\
\beta &= \frac{\left( \frac{i_{j+1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{b}} &= x \mathbf{b}_x + y \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
P_{\mathbf{s}} &= x \mathbf{s}_x + \beta y \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
&\forall \  P \in \text{P} \\ 
\end{split}
\end{align}

The effect of this transformation can be seen in the points represented in Fig. \ref{fig:Coordinates}, shown in scale for both coordinate systems. We can consider our points residing in the Cartesian coordinate system after the transformation.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set P of points, in both the original (top) and standard (bottom) coordinate systems, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

@@

Algorithms, such as the convex hull and the alpha shapes, solve this problem by discriminating which points of a set are part of its frontier.

A formal definition of alpha shapes is given at \cite{1983Edelsbrunnershape}. In the context of this work, however, gaining the intuition of the algorithm is more important, and thus the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.

The explanation, translated to our context, invites one to think of the plane where the points of P are represented as a soft material, such as Styrofoam. The points themselves are made of a harder material, such as rock, and occupy a fixed position in this plane. The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $. At the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.

@@

Discrete curvature estimation is an important task in image processing \cite{2010FleischmannNovel} for which no default definition exists.

@@

Thus, the set V of the vectors from each point of P to the next, shown in the Fig. \ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the average curvature of the discrete curve.

@@

We assume that, near the vector $ v_k $, the envelope defines a vertical displacement $ v_{k,y} = y_{j+1} - y_j $ and a horizontal displacement $ v_{k,x} = x_{j+1} - x_j $. Since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.

@@

The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}. This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.

The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally distant from each other.

@@

For calculation purposes, however, the formulation in Equation \ref{eq:Radius} poses a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $. In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero. The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Equation \ref{eq:Curvature}.

@@

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set $ E $ of the points belonging to the envelope if and only if exists one such circle defining a disc that does not contain any other point of P. Summarizing this, we have Equation \ref{eq:InEnvelope}:

@@

The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope. Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle as far as possible to the left or right of the first or last point, respectively. The discs defined by those circles can't contain any other point.

From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.

The procedure is formalized in algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.

@@

Recall from Section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Equation \ref{eq:Basis}. We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.

After converting E to the original coordinate system, the most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.

@@

The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task, for which there are no agreed-upon objective metrics. Recalling Fig. \ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the envelope, the carrier wave thus obtained is expected to be bounded between -1 and 1.

Figs. \ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted envelope.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spokenvoice}
\end{figure}

Those characteristics are in stark contrast with those of the singing voice in Fig. \ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown. There, the wave is almost periodic, with a stable waveform and envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The fast frequency variations in the signal didn't affect the envelope extracted.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

@@

In an effort to formalize the assessment of the quality of amplitude and frequency modulations of a signal, \cite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.

@@

More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
The set E of the points that belong to the envelope is a subset of P. We can thus write Equation \ref{eq:Req1}:

@@

This requirement doesn't imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.

@@

For any given frequency $ f $, both $ m_\mathbf{w}(f) $ and $ m_\mathbf{c}(f) $ are nonnegative. Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $ we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $ which is a more restrictive condition than the original: If $ \mathbf{w} $ has a finite frequency range, the magnitude of frequencies outside that range are zero, as are those of the same frequencies of the carrier $ \mathbf{c} $.

The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, generate a constant envelope $ \mathbf{e} $ of magnitude $ a $ and a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.

An example of such behaviour can be observed in Fig. \ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown. To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $. As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus, constant. Likewise, since $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we have $ c_i = a \cos(\omega i + \phi) / a = \cos(\omega i + \phi) \ \forall \ i, 0 \le i \le n - 1 $.

@@

Recall from Section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system. After the extraction, the inverse operation is performed on the points of the Envelope.

Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Equation \ref{eq:Basis}.

Given a constant $ \alpha > 0, \alpha \in \mathbb{R} $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $. Observing, from Equation \ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.

This guarantees that the same subset of points E will be identified as part of the envelope in both cases.
In the respective original basis system, applying the inverse change of basis, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert)  \quad \forall \  P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ for the scaled signal.

@@

In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off applicability in favor of mathematical formality.
In \cite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves. Fig. \ref{fig:RefEnvelopeTheory} illustrates this concept, for a family of sinusoids that share the same frequency and amplitude, and whose phases vary from 0 to $ 2 \pi $.

@@

In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal. Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/10RefEnvelope.pdf}
  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
  \label{fig:RefEnvelope}
\end{figure}

The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in Fig. \ref{fig:RefEnvelope}, besides being computationally expensive to calculate. However, arising from a mathematically sound background, it can serve as a reference to infer the quality of empirical methods.

@@

Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \cite[e.g.,][]{2018AssefModeling}.

Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms. Specifically, we compare the present method with the following approaches:

@@

In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \cite{1996Loughlinamplitude}, which states that the envelope of a periodic wave should be a straight line.

@@

For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable undershooting, as can be seen in Fig. \ref{fig:Comparison_piano}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

Besides, the traditional methods demanded careful choice of parameters. In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen. That is in contrast with our method, which organically tunes itself via the automatic choice of the circle's radius representing the average discrete curvature of the signal.

>>>>

\begin{abstract} 
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition. Nevertheless, a general approach to envelope detection does not exist. Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation. To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning. The method draws inspiration from geometric concepts: the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope. Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave. The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible. We provide visualizations of the envelope extracted for various real-world signals with diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes. An optimized implementation is available via the Python Package Index (PyPI), while interactive visualizations and source code are available from the accompanying GitHub repository.
\end{abstract}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00AlphaShape.pdf}
  \caption{Alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that does not contain any point of the set, is also shown.}
  \label{fig:AlphaShape}
\end{figure}

@@

While there is no general formal definition for the envelope of broadband signals \cite{2008HlawatschTime,2012HuEMD,2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, without intersecting with it.

@@

Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform do not behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempirical}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.

@@

Geometrically, we can think of these points as lying in a vector space with a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in representing those points in the Cartesian coordinate system, and for that we need a transformation that allows those points to be described in the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.

@@

In this way, we can consider both axes as having unit $ i $. Dividing them by $ i $, we obtain the effect of representing the points in the standard basis. The relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points in the original coordinate system.

Considering $ P_{\mathbf{b}} $, a point of the set P represented in the original coordinate system and $ P_{\mathbf{s}} $ the same point represented in the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:

\begin{align}\label{eq:Basis}
\begin{split}
P & = (x, y) \\
\beta &= \frac{\left( \frac{i_{j+1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{b}} &= x \ \mathbf{b}_x + y \ \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
P_{\mathbf{s}} &= x \ \mathbf{s}_x + \beta \ y \ \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
&\forall \  P \in \text{P} \\ 
\end{split}
\end{align}

The effect of this transformation can be seen in the points represented in Fig. \ref{fig:Coordinates}. Despite the two bases being different, the horizontal relationship is preserved, since one unit along the $ x $ axis in Cartesian coordinates corresponds to one unit of the sample index $ i $ in the original coordinate system.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set P of points, in both the original (top) and Cartesian (bottom) coordinate systems, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

@@

Algorithms such as the convex hull and the alpha shapes solve this problem by discriminating which points of a set are part of its frontier.

A formal definition of alpha shapes is given in \cite{1983Edelsbrunnershape}. In the context of this work, however, gaining intuition about the algorithm is more important, and for that the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.

The explanation, translated to our context, invites one to think of the plane where the points in the set P are represented as a soft material, such as Styrofoam. The points themselves are made of a harder material, such as rock, and occupy a fixed position in this plane. The alpha shapes algorithm can be seen as carving this plane from the outside of those points with a circular eraser of fixed radius $ \alpha $. At the end of the process, the remaining Styrofoam that could not be erased due to the impeding rocks is the alpha-shape of this set of points.

@@

Discrete curvature estimation is an important task in image processing \cite{2010FleischmannNovel}, for which no default definition exists.

@@

Thus, the set V of the vectors from each point of P to the next, shown in the Fig. \ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the curvature of the set.

@@

We assume that, near the vector $ v_k $, the envelope defines a vertical displacement $ v_{k,y} = y_{j+1} - y_j $ and a horizontal displacement $ v_{k,x} = x_{j+1} - x_j $. Since we do not yet know which points of P are part of the envelope, we cannot be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.

@@

The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}. This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when there is no vertical displacement, effectively turning the circle into a line.

% The radius is also directly proportional to $ v_{k,x} $, becoming larger in proportion to the horizontal distance between the adjacent points.

@@

For calculation purposes, however, the formulation in Equation \ref{eq:Radius} poses a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $. In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero. The average of the curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Equation \ref{eq:Curvature}.

@@

For all circles of center $ (x_o, y_o) $ and radius $ r $ that can be constructed to pass through any two points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set E of the points belonging to the envelope if and only if exists one such circle defining a disc that does not contain any other point of P. Equation \ref{eq:InEnvelope} summarizes this.

@@

The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope}, after noting that the first and last points of P will always be part of the envelope. Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing the first, leftmost point of P, by placing the center of the circle at $ (P_{0,x} - r, P_{0,y}) $. The disc defined by this circle cannot contain any other point of P, since all other points have abscissas greater than $ P_{0,x} $. The same is true for $ P_{m-1} $, the last and rightmost point of P.

From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that does not contain any point is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.

The procedure is formalized in algorithm \ref{FindEnvelope}. We do not provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.

@@

Recall from Section \ref{sec:mapping} that the coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Equation \ref{eq:Basis}. We could invert the transformation, but since we were careful to maintain a link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.

After converting E back to the original coordinate system, the most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.

@@

The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task \cite{2012HuEMD}, for which there are no agreed-upon objective metrics. Recalling Fig. \ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the extracted envelope, the carrier wave obtained is expected to be bounded between -1 and 1.

Figs. \ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present real-world discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier a male voice uttering the word “amazing”.}
  \label{fig:spokenvoice}
\end{figure}

Those characteristics are in stark contrast with those of the singing voice in Fig. \ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown. There, the wave is almost periodic, with a stable waveform and envelope. The superior frontier of the carrier wave in this case is approximately horizontal.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for an alto singer sustaining a steady note.}
  \label{fig:singer_voice}
\end{figure}

Fig. \ref{fig:Bend} shows the representation of a bend performed in an electric guitar. The fast frequency variations in the signal did not affect the envelope extracted, nor the carrier wave.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

@@

In an effort to formalize the assessment of the quality of the amplitude and frequency modulations of a signal, \cite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.

@@

More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
The set E of the points that belong to the envelope is a subset of P. We can thus write Equation \ref{eq:Req1}, with P represented in the original coordinate system:

@@

This requirement does not imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.

@@

For any given frequency $ f $, both $ m_\mathbf{w}(f) $ and $ m_\mathbf{c}(f) $ are nonnegative. Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $, we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $, which is a more restrictive condition than the original. If $ \mathbf{w} $ has a finite frequency range, the magnitude of the frequencies outside that range are zero, as are those of the carrier $ \mathbf{c} $.

The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a \in \mathbb{R}, \ a > 0  $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, generate a constant envelope $ \mathbf{e} $ of magnitude $ a $ and a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.

An example of such behaviour can be observed in Fig. \ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown. To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $ in the original coordinate system. As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus, constant. Likewise, since $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we have $ c_i = a \cos(\omega i + \phi) / a = \cos(\omega i + \phi) \ \forall \ i, 0 \le i \le n - 1 $.

@@

Recall from Section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system. After the extraction, the inverse operation is performed on the points of the envelope.

Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema. Represented in the Cartesian coordinate system, $ P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Equation \ref{eq:Basis}.

Given a constant $ \alpha \in \mathbb{R}, \ \alpha > 0 $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $. Observing, from Equation \ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \} $, with $ P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $, is equal to the original set P.

This guarantees that the same subset of points E will be identified as part of the envelope in both cases.
In their respective original basis system, applying the inverse transformation, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert) = (i_j, |w_{i_j}|) \quad \forall \ P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) = (i_j, |\alpha w_{i_j}|)  \quad \forall \ P^\prime \in \text{E}^\prime $ for the scaled signal.

@@

% In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, 
In \cite{1992BruceCurves,2019Jiaempirical}, the envelope is defined as the boundary of the region filled by a family of curves. Fig. \ref{fig:RefEnvelopeTheory} illustrates this concept, for a family of sinusoids that share the same frequency and amplitude, and whose phases vary from 0 to $ 2 \pi $.

@@

In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal. Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
Fig. \ref{fig:RefEnvelope} shows part of the reference envelope for the sound of a tom drum.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/10RefEnvelope.pdf}
  \caption{Part of the reference envelope obtained by shifting the original signal horizontally, for the sound of a tom drum.}
  \label{fig:RefEnvelope}
\end{figure}

The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in Fig. \ref{fig:RefEnvelope}. However, those segments agree with the general contour of the wave, and overestimate and underestimate the envelope in equal measures.

Although lacking in applicability, both given its discontinuous shape and the high computational cost, an envelope constructed in this way arises from a mathematically sound background, and can serve as a reference to infer the quality of more practical methods.

@@

Direct comparison with many recent algorithms is made difficult by the unavailability of digital implementations of such works, some of whice where designed to process analog signals \cite[e.g.][]{2018AssefModeling}.

Nevertheless, insight can be gained from comparing the method here proposed with some of the most common envelope extraction algorithms. Specifically, we compare the present method with the following approaches:

@@

Fig. \ref{fig:Comparison_sinusoid} presents the comparison in the case of a simple sinusoid. 
The present algorithm and the Hilbert transform approach comply with the third condition proposed by \cite{1996Loughlinamplitude}, which states that the envelope of a periodic wave should be a straight line.

@@

For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach undershoots the original signal, as can be seen in Fig. \ref{fig:Comparison_piano}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the sound of key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

In Fig. \ref{fig:Comparison_soprano} a soprano singer sustaining a steady note is shown. All envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

Besides, the traditional methods demanded careful choice of parameters. In the case of the Hilbert transform, post-processing in the form of filtering was needed, and an appropriate cut-off frequency had to be chosen. That is in contrast with our method, which organically tunes itself via the automatic choice of the circle's radius representing the average discrete curvature of the signal.

Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of Section \ref{sec:quality}.
